---
title: "airport-network"
author: "Thomas Gredig"
date: "3/13/2019"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```

# Airport Network

The data is loaded from [OpenFlights.Org](https://openflights.org/data.html).

The following data is available:

```{r}
path.source = 'data'
file.list = dir(path.source, pattern='dat$')
table(file.list)
```

## Airports

```{r}
d = read.csv(file.path(path.source, 'airports.dat'), header=FALSE)
levels(factor(d$V14))
d[,c(13,14)] <- NULL
kable(d[1:5,])
```

```{r}
summary(d)
kable(d[grep('Long Beach',d$V2),])
```


## Airports in Switzerland

Finding all the airports in Switzerland

```{r}
kable(d[grep('Switzerland',d$V4),])
```

# Routes

Let us look at the routes database

```{r}
dr = read.csv(file.path(path.source, 'routes.dat'), header=FALSE,
              stringsAsFactors = FALSE)
summary(dr)
kable(dr[1:5,])
```

Now also load the airlines and then find all "Swiss" airlines.

```{r}
da  = read.csv(file.path(path.source, 'airlines.dat'), header=FALSE,
               stringsAsFactors = FALSE)
da[grep('Swi',da$V2),]
```

Now, that we know *4559* is Swiss Airlines, we can find the connections:

```{r}
dr$V2 = as.numeric(dr$V2)
dr[which(dr$V2==4559),]
```


There are `r length(which(dr$V2==4559))` flights with "Swiss".

# Graphs

We are following the tutorial on [Network Analysis in R with iGraph](http://kateto.net/networks-r-igraph). We will use the `graph` function, which uses `edges`

```{r}
library(igraph)
g1 <- graph( edges=c('a','b','a','c','x','d','c','d'),  directed=F ) 
plot(g1)
dr[which(dr$V2==4559),] -> q
q = q[1:200,]
g1 <- graph( edges=c(rbind(q$V3, q$V5)) , directed=F)
plot(g1,vertex.label.cex=.5,edge.color="orange", vertex.size=15, layout=layout_nicely,vertex.color="gray90" )
```

Find the names of some of the airports

```{r}
d[which(d$V5 %in% c('BCN','BRU','BCN','ARN','DAR','NBO')),]
```

# Appendix

This was generated in `r getwd()` on `r Sys.Date()`.